---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r loadData, echo=TRUE}
require(dplyr)
activity <-read.csv("activity.csv",header=TRUE)
act_tb<-tbl_df(activity)
head(act_tb)
```

## What is mean total number of steps taken per day?

First the data is grouped by day, and then the average number of steps is computed for each day

```{r numberSteps}
by_date<-group_by(act_tb,date ) %>%
summarize(pasos=sum(steps))    
hist(by_date$pasos,breaks=20, main="Histogram of steps taken per day", col="blue", xlab="Number of Steps")
```

The average daily steps is `r format(mean(by_date$pasos,na.rm=TRUE))` and the median is `r format(median(by_date$pasos,na.rm=TRUE))`

## What is the average daily activity pattern? 
For the activity pattern, the data is grouped by interval of time (HHMM) 

``` {r activityPattern}
by_interval<-group_by(act_tb,interval) %>% summarize(media=mean(steps,na.rm=TRUE))

plot(by_interval$interval,by_interval$media,type='l',main="",xlab="Time of the day (HHMM)",
     ylab="Average steps taken")
     
```
``` {r maxActivity}
temp<-filter(by_interval,media==max(media))

```
On average across all the days in the dataset, the interval that contains the maximum number of steps is `r temp$interval` with an average steps of `r format(temp$media)`


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
